### 2025-05-20 analysis/paper/paper1.qmd
updates

---
title: "Test"
format: html
bibliography: references.bib
---
这是 Wadell 的研究引用示例 [@Wadell1935]